
\chapter{Results}
This chapter is dedicated to the results of model testing on users, as well as the inpact of features and hyperparameter choices on its performance. 


\subsubsection{Methodology}
\myworries{Word this nicely}
The results below were calcutaled as an average of each models perfomance on a sequance of characters equal to the length on which the model was trained.
This means that a some models perfomed inference on sequences of lenght 20, while others 40. While it was possible to compare scores based on an equal input lenght,
it would require the setting another threshold, on how many of the input subsequnces need to be classified possitivly for the whole sequance to be classified as possitive. 
Such threshold was chosen for the aplication, however, we believe that adding another parameter to these results would only make them less interpretable.

\myworries{Tutaj macierz pomyłem między wszyskimi modelami, ostateczne wyniki, najlepsze parametry etc Dezycja czy w sekcji -- moze byc ale jak ją nazwać}

\subsubsection{Confusion matrix for all users}

\subsubsection{Roc curves}
\myworries{Points in roc space}

\subsubsection{Equal error rates}
\myworries{EER for each user, the threshold neccessary}

\section{Input features}
This section will compare the impact of input encodings on the final model performance.

\subsection{Input sequance lenght}

\subsection{Egde data encoding}

\subsection{Character encoding}

\subsection{Accelerometer Data}
TLDR doesnt work, comparison for 2 users only - best performing on normal input
Short discussion

\section{Training and model hyperparameters}

\subsection{Class imbalance}
\myworries{Discuss how positive/negative example ratio had an impact on precission/recall. Maybe note that neg examples were sampled with an offset}

\subsection{Model size}
\myworries{Num of conv layer, layer size etc bigger wasnt better}


\section{Testing model on users}

\subsection{TODO title}
\myworries{ TODO: What if we did classify the user on all their input (100 chars). Results would be a n users by n users matrix of 1's and 0's}

\subsection{Multiclass model}
\myworries{Maybe}


\subsection{Cross-smartphone user validation}
TODO: what happens if two users train on smartphones that are not their own? What happens, if they cross-use their original model on another phone?

\section{Discussion}
TODO: discuss the findings.