
\chapter{Conlusion}

In summary, the approach used in this project and study requires further investigation to be effectively evaluated. The effects of such an approach seem promising but are currently behind the state-of-the-art models. A larger study would also be a source of validation or rebuttal to many hypotheses posed in the discussion chapter about the possible sources of error in the model's performance on some specific users. A larger dataset specifically could also benefit the scientific community at large, if the users would agree to release it under an open-source license. Data gathered in this study can still be considered valuable and can be used for small-scale testing, it may however be of lesser quality for some specific users. Alternatively, such data can be used to demonstrate some of the common actions that the users could perform, which would negatively impact the method's performance.

It is worth highlighting that for many users, the model performed well. Further refinement or combining the model with other neural network types, feature engineering, or modification to the data gathering process could also lead to improved results.

The team considers all the project goals achieved. The application used for data gathering is functional and -- with a few changes and fixes -- could be a basis for an application to be used in possible further research. The application can also be used as a demonstration tool for the model, allowing users to test the inference functionality in real-time, if connected to the server. The server works well for a small number of users, as was assumed at the start of the project. It can be modified to work in the cloud environment if needed. For budget reasons, this was not done for this project. The models created are functional, as discussed in Chapter 6. Data gathering operation was mostly successful, with some problems related to specific users. In further studies, these problems could be remedied in ways described in Chapter 6. Thus, the team considers the project successful in demonstrating that the proposed approach has merit and warrants further exploration.